---
title: "Nerd Nite"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(lubridate)
library(scales)
library(plotly)
```

Load in the data and create two new columns, one for if the Meet was the Olympics and one for if the Meet was World Champs.

```{r data load, warning = FALSE}
setwd("C:/Users/Orly Olbum/OneDrive - Inpleo/Desktop/Misc/Nerd Nite Feb 2022")
world = read.csv("WRData.csv")

# see which ones are Olympics/Worlds vs. not
world = world %>%
  mutate(Olympics = ifelse(grepl("Olympic Games", Meet, TRUE), TRUE, FALSE),
         Worlds = ifelse(grepl("World Champs", Meet, TRUE), TRUE, FALSE))

# add column for stroke
world = world %>%
  mutate(Stroke = gsub("[^a-zA-Z]", "", Event))

# get year from date (first format date)
world = world %>%
  mutate(Date = mdy(Date),
         Year = year(Date))
```

Summarize the data:
How many world records are there per event?

```{r, warning = FALSE}
world %>%
  group_by(Event, Gender) %>%
  summarise(Count = length(Event))
```

World records by country

```{r}
world %>%
  group_by(Swimmer_Country) %>%
  summarise(Count = length(Event)) %>%
  arrange(desc(Count))
```

World records by stroke

```{r}
world %>%
  group_by(Stroke) %>%
  summarise(Count = length(Event)) %>%
  arrange(desc(Count))
```

Top 10 swimmers M/F by number of world records

```{r}
# women
world %>%
  filter(Gender == "F") %>%
  group_by(Swimmer, Swimmer_Country) %>%
  summarise(Count = length(Event)) %>%
  arrange(desc(Count)) %>%
  head(10)
```

```{r}
# men
world %>%
  filter(Gender == "M") %>%
  group_by(Swimmer, Swimmer_Country) %>%
  summarise(Count = length(Event)) %>%
  arrange(desc(Count)) %>%
  head(10)
```

How many world records were broken in each Olympic games?

```{r}
world %>%
  filter(Olympics == TRUE) %>%
  group_by(Meet) %>%
  summarise(Count = length(Event))
```

```{r}
# for bar plot of WRs by Olympic Year
olympic_wrs = world %>%
  filter(Olympics == TRUE) %>%
  group_by(Year) %>%
  summarise(Count = length(Event))

# plot bar chart
plot_ly(
  data = olympic_wrs,
  x = ~Year,
  y = ~Count,
  type = "bar",
  text = ~Count,
  name = "",
  hovertemplate = paste("%{x}", "<br>%{text}")
) %>%
  layout(
    title = "WRs by Olympic Games",
    xaxis = list(
      title = "Olympic Games"
    ),
    yaxis = list(
      title = "World Records"
    )
  )
```

How many world records were broken in each World champs?

```{r}
world %>%
  filter(Worlds == TRUE) %>%
  group_by(Meet) %>%
  summarise(Count = length(Event))
```

How many world records were broken by year overall?

```{r}
records_year = world %>%
  group_by(Year) %>%
  summarise(Count = length(Event))

# plot bar chart
plot_ly(
  data = records_year,
  x = ~Year,
  y = ~Count,
  type = "bar",
  text = ~Count,
  name = "",
  hovertemplate = paste("%{x}", "<br>%{text}")
) %>%
  layout(
    title = "WRs by Year",
    xaxis = list(
      title = "Year"
    ),
    yaxis = list(
      title = "World Records"
    )
  )
```

How many world records were broken in/not in an Olympics or Worlds?

```{r}
# in Olympics or Worlds
world %>%
  filter(Olympics == TRUE | Worlds == TRUE) %>%
  summarise(Count = length(Event))

# not in Olympics or Worlds
world %>%
  filter(Olympics == FALSE & Worlds == FALSE) %>%
  summarise(Count = length(Event))
```

How many world records were broken by GDR women between 1970 and 1980 (height of doping)? What percentage of all world records?

```{r}
gdrw_wrs = world %>%
  filter(Swimmer_Country == "GDR" &
           Gender == "F" &
           Year >= 1970 & Year <= 1980) %>%
  summarise(Count = length(Event))

all_wrs = nrow(world)

gdrw_wrs[1,]
percent(gdrw_wrs[1,]/all_wrs)
```

The textile suits came out in  2008 and were banned at the end of 2009. How many world records were broken in those 2 years? What percentage of all WRs were in this timeframe?

```{r}
tech_wrs = world %>%
  filter(Year == 2008 | Year == 2009) %>%
  summarise(Count = length(Event))

all_wrs = nrow(world)

tech_wrs[1,]
percent(tech_wrs[1,]/all_wrs)
```

The GDR women make up almost as many world records as the LZR records of 2008-2009 - which is perhaps more evidence suggesting the suits shouldn't have been allowed in the first place. It may be up for discussion if they were considered technical doping, but there was no doubt about the East German's doping.

Plot World Record times over Time for each freestyle event, colored by men and women.

```{r}
# line for years of stuff
vline = function(x = 0, color = "green") {
  list(
    type = "line",
    y0 = 0, y1 = 1, yref = "paper",
    x0 = x, x1 = x,
    line = list(color = color, dash = "dot")
  )
}
```

```{r, warning = FALSE, fig.width = 10}
# 50 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "50 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "50 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "50 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

```{r, warning = FALSE, fig.width = 10}
# 100 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "100 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "100 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "100 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

```{r, warning = FALSE, fig.width = 10}
# 200 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "200 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "200 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "200 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

```{r, warning = FALSE, fig.width = 10}
# 400 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "400 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "400 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "400 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

```{r, warning = FALSE, fig.width = 10}
# 800 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "800 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "800 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "800 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

```{r, warning = FALSE, fig.width = 10}
# 1500 free
plot_ly() %>%
  add_markers(
    data = world[world$Event == "1500 FR" & world$Gender == "F",],
    x = ~Date, y = ~Time, name = "Female"
  ) %>%
  add_markers(
    data = world[world$Event == "1500 FR" & world$Gender == "M",],
    x = ~Date, y = ~Time, name = "Male"
  ) %>%
  layout(
    title = "1500 Free WRs over Time",
    xaxis = list(
      title = "Date"
    ),
    yaxis = list(
      title = "Seconds"
    ),
    shapes = list(
      vline("1967-01-01"),
      vline("1976-01-01"),
      vline("2008-01-01")
    )
  )
```

Replicate some stat analysis from https://sites.northwestern.edu/elannesscohn/2019/07/03/disruptive-sports-tech-exploratory-analysis-of-swim-world-record-data/ modeling the freestyle (non-50) events for men and women before and after the goggles were brought to the Olympics in 1976.

```{r}
# prep data for the modeling
# want only freestyle and not the 50
free = world %>%
  filter(Event == "100 FR" | 
           Event == "200 FR" | 
           Event == "400 FR" | 
           Event == "800 FR" | 
           Event == "1500 FR") %>%
  mutate(Goggles = ifelse(Year < 1976, TRUE, FALSE))

# set up models
mod1 = lm(Time ~ Gender + Goggles, data = free)
summary(mod1)

mod2 = lm(Time ~ Gender + Goggles + Date, data = free)
summary(mod2)
```







